### 2015-02-10 2015-eamt-assim/eng-kaz/tables-updated.tex
move dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58814 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

%\date{}							% Activate to display a given date or no date

\begin{document}
%\section{}
%\subsection{}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&40.48 +/- 45.30&30.16 +/- 35.86&38.10 +/- 46.05&35.33 +/- 38.68\\
20\%&54.63 +/- 27.53&39.19 +/- 20.37&24.79 +/- 19.16&48.78 +/- 32.65\\
30\%&23.87 +/- 24.95&35.13 +/- 30.18&30.41 +/- 34.73&43.07 +/- 32.71\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&31.43&27.50&34.15&30.43\\
20\%&51.72&36.23&22.54&43.84\\
30\%&21.85&32.58&25.56&35.63\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&40.48 +/- 45.30&42.06 +/- 38.70&42.86 +/- 44.42&39.33 +/- 37.38\\
20\%&63.89 +/- 28.87&49.78 +/- 19.89&36.01 +/- 24.61&48.78 +/- 32.65\\
30\%&32.08 +/- 27.38&49.42 +/- 25.61&45.01 +/- 37.62&47.04 +/- 33.64\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&31.43&40.00&39.02&34.78\\
20\%&60.34&47.83&33.80&43.84\\
30\%&31.09&44.94&36.56&41.38\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0:47 +/- 0:35&1:47 +/- 1:36&1:34 +/- 1:03&3:18 +/- 2:35\\
20\%&1:13 +/- 1:02&2:24 +/- 2:21&2:19 +/- 2:07&1:31 +/- 1:21\\
30\%&4:20 +/- 11:28&2:01 +/- 1:49&3:21 +/- 4:18&2:00 +/- 1:30\\
\hline
\end{tabular}
\caption {Time taken to fill the gaps in one sentence (m:s)} \label{tab:title} 
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{1}{c}|}
\hline
Gap percentage & Number of holes\\
\hline
10\%&22\\
20\%&43\\
30\%&52\\
\hline
\end{tabular}
\caption {Total number of holes to be filled by one evaluator.} \label{tab:title}
\end{table}

\begin{table}
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0.812&0.000&0.516&0.143\\
20\%&0.638&0.282&0.425&0.507\\
30\%&0.521&0.348&0.795&0.413\\
\hline
\end{tabular}
\caption {Krippendorff Alpha measure} \label{tab:title}
\end{table}

\end{document}