### 2015-02-19 2015-eamt-kazkir/2015-eamt-kazkir.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59025 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
% eamt12.tex)
%
% Contact: eamt2015@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt15}
\usepackage{times}
\usepackage{latexsym}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}
\usepackage{url}

\newcommand{\confname}{EAMT 2015}
\newcommand{\website}{http://www.eamt2015.org/}
\newcommand{\contactname}{the conference chairs (Felipe
  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
\newcommand{\conffilename}{eamt15}
\newcommand{\downloadsite}{http://www.eamt2015.org/}
\newcommand{\paperlength}{$8$ (eight)}
\newcommand{\shortpaperlength}{$4$ (four)}

\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}

\title{A comparison of approaches to Kazakh to Kyrgyz machine translation}

\author{
  Jonathan North Washington\\
  Departments of Linguistics\\
  and Central Eurasian Studies\\
  Indiana University\\
  Bloomington, Indiana 47405 USA\\
  \texttt{jonwashi@indiana.edu}  \And
  \textbf{Francis M. Tyers}\\
  HSL-fakultetet\\ 
  UiT Norgga \'{a}rktala\v{s} universitehta \\
  9017 Romsa, Norway \\
  {\tt francis.tyers@uit.no}
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}

This paper presents a comparison of approaches to machine translation
from Kazakh to Kyrgyz.
The paper will be laid out as follows: Section~\ref{sec:prev}\ gives a short review of some previous work in the area of Turkic--Turkic language machine translation; Section~\ref{sec:lang} introduces Kazakh and Kyrgyz and compares their grammar;
Section~\ref{sec:sys}\ describes the systems we built and the tools used to construct them;
Section~\ref{sec:eval}\ gives an evaluation of the systems; and
finally Section~\ref{sec:conc}\ describes our aims for future work and some
concluding remarks.

\section{Previous work}
\label{sec:prev}

Within the Apertium project, work on several MT systems between Turkic languages has been started (Kazakh--Tatar, Tatar--Bashkir, Turkish--Kyrgyz, and Azeri--Turkish).  There are two published works describing these systems, the first is \cite{tyerswashingtonsalimzyanbattalov12} which describes a Tatar--Bashkir machine translation system; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage. The second is \cite{tyerswashingtonsalimzyan13} which describes a high-coverage Kazakh--Tatar system.

Besides these systems, several previous works on making machine translation systems between Turkic languages 
exist, although to our knowledge none are publicly available except for the Turkish--Azerbaijani pair available through Google Translate.\footnote{\url{http://translate.google.com}}
Some MT systems have been reported that translate between Turkish and other Turkic languages, 
including Turkish--Crimean Tatar \citep{altintas01},
Turkish--Azerbaijani \citep{hamzaoglu93}, Turkish--Tatar \citep{suleymanov08}, and
Turkish--Turkmen \citep{tantug07}, though none of these have been released to a public audience.

\section{Kazakh and Kyrgyz}

\section{Approaches}

\subsection{Rule-based machine translation}

\cite{washingtonsalimzyantyers14}
\cite{washingtonipasovtyers12}

\subsection{Corpus-based machine translation}

We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \cite{zens2002}) and the second on character-based machine translation (CBMT, \cite{nakov2012}; \cite{TiedemannEAMT2009}).

\subsection{Corpus}

\subsubsection{Phrase-based}

\subsubsection{Character-based }

\cite{nakov2012} present a method of statistical machine translation on the character level between related languages that takes advantage of phrase-based machine translation architecture. 
%The method relies on preprocessing the input and output by inserting spaces in between the characters of words, for example the string `m\"{a} meen Helsinkiin' would become `m \"{a} \$ m e e n \$ H e l s i n k i i n' with a unigram model, or `m\"{a} \"{a}\$ \$m me ee en n\$ \$H He el ls si in nk ki ii in' with a bigram model.

%After preprocessing, the corpora are processed as with the phrase-based system, with the difference that the language model order is increased from 5 to 10-grams.

\section{Evaluation}

\subsection{Coverage}

\begin{table}
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    \textbf{Corpus} & \textbf{Tokens} & \textbf{RBMT} & \textbf{PBSMT} & \textbf{CBMT} \\
    \hline
    NT              & 144,946      &      &          &       \\
    Akorda          & 382,716       &      &          &       \\
    Wikipedia       & 27,802,493       &      &          &       \\
    \hline
  \end{tabular}
  \caption{Vocabulary coverage for the three systems}
  \label{table:coverage}
\end{table}

\subsection{Translation quality}

\section{Future work}

\section{Conclusions}

\section*{Acknowledgements}

\emph{Anonymised for review}

\bibliographystyle{apalike}
\bibliography{2015-eamt-kazkir}

\end{document}