### 2015-03-04 engkaz-eamt-2015/engkaz-eamt-2015.tex
making changew
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59212 72bbbca6-d526-0410-a7d9-f06f51895060

% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
% eamt12.tex)
%
% Contact: eamt2015@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt15}
\usepackage{times}
\usepackage{latexsym}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%
\usepackage{url}

\usepackage{multirow}

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}
\usepackage{latexsym}
%\usepackage[small,bf]{caption}
\usepackage[bf]{caption}
\usepackage{xltxtra}
\usepackage{times}
\usepackage{fontspec}
\usepackage{footnote}
\usepackage{natbib}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\usepackage{listings}
\usepackage{microtype}

\newfontfeature{IPA}{+mgrk}
%\setromanfont[IPA]{FreeSerif}
%\setromanfont[Scale=0.9]{Times New Roman}
\setromanfont{Times New Roman}
\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
\newfontfamily\qipb[IPA,Scale=MatchLowercase]{Junicode}
\setmonofont[Scale=0.7]{DejaVu Sans Mono}
\newfontfamily\smallertt[Scale=0.55]{DejaVu Sans Mono}
\newfontfamily\smallerrm[Scale=0.85]{Times New Roman}

%\fontspec[FakeBold=2.5]{Times New Roman}
%\DeclareFontShape{EU1}{TimesNewRoman(0)}{m}{sc}{<->ssub * TimesNewRoman(1)/m/sc}{}

\newcommand{\confname}{EAMT 2015}
\newcommand{\website}{http://www.eamt2015.org/}
\newcommand{\contactname}{the conference chairs (Felipe
  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
\newcommand{\conffilename}{eamt15}
\newcommand{\downloadsite}{http://www.eamt2015.org/}
\newcommand{\paperlength}{$8$ (eight)}
\newcommand{\shortpaperlength}{$4$ (four)}

\newcommand{\com}[1]{\marginpar{\scriptsize #1}}

\title{A free/open-source machine translation system for English to Kazakh}

% Authors:
% Aida, Mikel, Fran, Inari, ... ?

\date{}

\begin{document}
\maketitle

\begin{abstract}
This paper presents the current state of development a shallow-transfer rule-based machine translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
between the two languages are presented, and it is showed how the MT 
system was designed to tackle these challenges. We show an evaluation of system coverage and 
translation quality and outline and future work.
\end{abstract}

\section{Introduction}

This paper presents a shallow-transfer rule-based machine translation (MT) system from English 
(a West Germanic language) to Kazakh (a Turkic language)
The complex agglutinative morphology of Kazakh 
is in contrast with the simple morphology of English.
Turkic language morphology shows clear morphotactics (ordering of morphemes), but with 
complex phonological changes (morphophonology) to due to morpheme contact (vowel harmony, 
sonorisation, etc.) many of which are explicitly represented in writing.

On the other hand, there are many differences between the syntax of Turkic languages 
and English: subject–object–verb order (vs.\ subject–verb–object in English), use of 
postpositions (vs.\ prepositions in English), head-final syntax with modifiers and specifiers 
always preceding the modified/specified (vs.\ following in English), overt case marking allowing 
for a rather free ordering of arguments (vs.\ a more fixed order in English), lack of definite 
articles (extensively used in English), verbal-noun-centered structures where English uses modal 
verbs (\emph{must}, \emph{have to}, \emph{want to}) or verbal-noun or verbal-adjective-centred constructions where 
English has subordinate clauses using finite verbs with relatives or subordinating 
conjunctions (\emph{the book which I read}, \emph{the place where I saw him}, \emph{before he came}), lack of a parallel 
of the English lexical verb \emph{to have}, etc. For an account (in Russian) of 
syntactic differences between English and Kazakh, see \cite{pecherskih2012}.

We currently build rule-based machine translation system, because for statistical machine translation 
system we need large sentence-aligned parallel corpora. There are no freely available, large parallel 
corpora for the English-Kazakh pair. English-Kazakh statistical MT system was built by Google,\footnote{\url{http://translate.google.com}} 
other work in this language pair was made by building a small parallel corpus and using Moses \citep{assylbekov14}, 
and existing commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based. Rule-based machine translation, in which developers 
write up dictionaries and grammatical rules that are applied by an engine, in some 
unusual case will produce better translation.

This paper describes work in progress of development of machine translation system 
for English-to-Kazakh, which developed by using Apertium free/open-source machine 
translation platform \citep{forcada11}. %(http://www.apertium.org).

\section{The Apertium platform}

\begin{figure*}[htbp]
\begin{center}
 \includegraphics[width=0.8\textwidth]{architecture.pdf}
\end{center}
\caption{The pipeline architecture of the Apertium system.}
\label{fig:modules}
%\vspace{-1em}
\end{figure*}

Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
has been developed since 2005 at the Universitat d'Alacant. At first, it was initially aimed at 
translating texts between closely related languages, then it was extended to deal with less-related 
languages. This platform has next components: machine translation engine, developer's tools, and linguistic 
data for an increasing number of language pairs and they are licensed under the 
free/open-source GNU General Public License (GPL, versions 2 and 3).

\begin{description}
\item[De-formatter:] Separates the text to be translated from the formatting tags.  Formatting tags are 
  encapsulated in brackets so they are treated as ``superblanks'' that are placed between words in 
  such a way that the remaining modules see them as regular blanks.  
\item[Morphological analyser:] For each surface form the morphological analyser generates one or more 
  lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), 
  and inflection information. 
%  In this example each source form has been analysed as one or more lexical forms: ``I'' is analysed 
%  into ``I'', where lexical category is subject pronoun (\texttt{<prn><subj>}) and first person (\texttt{<p1>}), could be 
%  masculine or feminine (\texttt{<mf>}), singular (\texttt{<sg>}); ``played'' is analysed into lexical verb (\texttt{<vblex>}) 
%  ``play'' with past simple tense (\texttt{<past>}) or it could be past participle (\texttt{<pp>}), so it has two 
%  analyses. Each analysis of word is separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical 
%  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
\item[Part-of-speech (POS) tagger:] This module chooses one of lexical forms of ambiguous word. As can be 
  seen from the previous example, the morphological analyser could deliver more than one lexical form and 
  choosing wrong one could produce errors in translation. POS tagger is based on a statistical model, which 
  based on hidden Markov models and it has been trained on source language texts, which processes the 
  result of the application of  on constraint-grammar rules \citep{karlsson95}, which are used to discard some 
  analyses using rules based on context. 
%For example, consider the morphological analysis of word ``book'':
%\begin{verbatim}
%^book/book<n><sg>
%/book<vblex><inf>
%/book<vblex><pres>$
%\end{verbatim}
%  This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. 
After  this module, all words have only one morphological analysis.
\item[Lexical transfer:] It reads each lexical form of source language and delivers a corresponding target language 
  lexical form. This module uses a bilingual dictionary. 
%For the English-Kazakh language pair, each lexical form of 
%  English word is translated into Kazakh as follows:
%\begin{verbatim}
%^I<prn><subj><p1><mf><sg>/мен<prn><pers><subj><p1><mf><sg>$ 
%^play<vblex><past>/ойна<v><tv><past>$
%\end{verbatim}
 Multiword units can be translated as a single word.
\item[ Lexical selection:] It uses rules for lexical words, which have many translations, to select one of the translations 
  in the target language according to context. The system is described in \cite{tyers12a}.
\item [Structural transfer:] This module uses pattern matching to identify sequences of lexical forms  (phrases or 
  segments), which need syntactical processing to translate grammatical differences between two languages (handling of 
  number, gender, etc.). It uses  files with rules, which specify  the syntactic transformation such as word reorderings, 
  lexical changes such as changes in prepositions and agreement between target language lexical forms. %Transfer rules can 
%  produce new sequences for the  target language, for instance, preposition-noun rule is used to build a sequence, where for 
%  noun is chosen right case, which depends on preposition.
%\begin{verbatim}
%in garden  --  ^бақша<n><sg><PXD><loc>$
% \end{verbatim} 
\item[Morphological generator:] Delivers the sequence of target-language lexical forms, produced by the structural transfer, 
  to a corresponding sequence of target language surface forms. The morphological generator executes a finite-state transducer 
  generated by compiling a morphological dictionary for the target language.     
  For Kazakh language a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
\item[ Post-generator:] Performs some minor orthographical operations in the target language (for instance, it generates the 
  English form cannot from can and not). This module is generated from file with rules which are very   similar in format to dictionary files. 
\item[ Reformatter:] It places format tags back into the text so that its format is preserved.
\end{description}

\section{Linguistic data}

Apertium uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer and
lexical selection rules. % and rules for part-of-speech tagging.

\subsection{Dictionaries}

Dictionaries are used in lexical processing and morphological analysis, there are use two types: morphological (monolingual) and bilingual.

\subsubsection{Monolingual dictionaries}

English dictionary uses to determine lexical forms of each surface forms, it contains of list of all lexical units, definition 
of source-language alphabet, grammatical categories, which is used to specify lexical forms, for example, like noun, verb, 
gender, etc., each categories have paradigms, they describes groups of correspondences between parts of surface forms and lexical forms.

For Kazakh side, is used morphological transducers, which are based on the Helsinki Finite State Toolkit \citep{hfst/2011}. It 
uses formalisms lexc for defining lexicons, twol formalism for morphophonological rules. Lexc formalism defines different 
word classes and subclasses. And twol determines phonological transformation like vowel harmony, desonorisation. For example, 
depend on preceding vowel, plural form \emph{-LAr} could become -тар or -тер depending on vowel harmony: kitap+tar, but mektep+ter.

\subsubsection{Bilingual dictionary}

Bilingual dictionary provides correspondences between lexical forms of source and target lanugages. Each entry 
consists of English word on the left side with part of speech and corresponding Kazakh word on the right side. 
Some entries could be ambiguous by part of speech, or just have many translations, which will be chosen by 
lexical-selection rules depending on context. This dictionary currently contains 13,130 stems \citep{sundetova13a}.

\subsection{Rules}

\subsection{Disambiguation rules}

This kind of rules solves the problem of part-of-speech (POS) ambiguity by Constraint Grammar (CG, \cite{karlsson95}). The 
output of morphological analyser could be ambiguous, and in this case CG rules choose correct analysis from 
multiple analyses. From English to Kazakh POS ambiguity could be only in English side, but also some choice also
depend on Kazakh grammar.

\subsection{Lexicon selection rules }

Lexical-selection rules choose one of the alternative translations of one word depend on the context. The rule 
formalism is described in \cite{tyers12a}. Alternative translations are defined in bilingual dictionary by 
multiple entries of one word. For exmaple, adjective ``beautiful'' could be translated as ``сұлу'', if 
following noun is subject: ``beautiful girl'' -- ``сұлу қыз''; or ``әдемі'', ``көркем'', if the noun means 
place or: ``beautiful mountains'' -- ``әдемі таулар''.  Table~\ref{table:lexsel} demonstrates some 
lexical-selection rules:

\begin{table*}
 \centering
 \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{Word} & \textbf{Translation} & \textbf{Condition} & \textbf{Example} \\
    \hline
    \multirow{4}{*}{residence} & \multirow{2}{*}{мекен}      & \multirow{2}{*}{--} & I live in a residence. \\
                               &                             &                     & (Мен) мекенде өмір сүремін. \\
                               & \multirow{2}{*}{резиденция} & \multirow{2}{*}{\_ of \_ president} & I see the residence of the president \\ 
                               &                             &                                     & (Мен) президенттің резиденциясын көремін \\
    \hline
    \multirow{4}{*}{boot} & \multirow{2}{*}{бәтеңке} & -- & I bought boots. \\
                          &                          &     &  Мен бәтеңкелерді сатып алдым \\

                          & \multirow{2}{*}{жүксалғыш} & \_ of \_ car &  He opened the boot of the car  \\ 
                          &           &              & Ол машинаның жүксалғышын ашты. \\
    \hline
    \multirow{4}{*}{anything} & \multirow{2}{*}{бір нәрсе} & -- & I see anything \\ 
                              &           &    & Мен бір нәрсе көремін \\
                              & \multirow{2}{*}{еш нәрсе} & not \_[verb] & I do not do anything \\ 
                              &          &             &  Мен еш нәрсе істемеймін. \\

    \hline
 \end{tabular}
  \caption{Example lexical-selection rules}
  \label{table:lexsel}
\end{table*}

\subsection{Transfer rules}

For English–Kazakh system this stage uses three stages to improve the granularity of structural-transfer 
rules \citep{sundetova13b}:

\begin{itemize}
\item A first round of transformations (``chunker'') detects source language (SL) LF patterns and generates the 
  corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc. 
\item The second round (``interchunk'') reads patterns of chunks and produces a new sequence of chunks. This is the 
  module where one can attempt to perform some longer-range reordering operations, interchunk agreement, case selection, etc. 
\item The third round (``postchunk'') transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level 
  tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is 
  transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
\end{itemize}

The structural transfer module in Apertium processes the stream of source-language lexical form -- target-language lexical 
form pairs (SL LF–TL LF pairs) and transforms it into a new sequence of TL LFs after a series of structural transfer 
operations specified in a set of rules: reordering, elimination or insertion of TL LFs, agreement, etc.

This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision ..., ). English to Kazakh 
chunker rules, interchunk rules and additional clean-up stage will be described in detail next 3 sections.

\subsubsection{Chunker}

In the first round of structural transfer rules do some local operations to cut sentences into chunks, like short noun 
phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and 
postpositional phrases in Kazakh).
Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much 
as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the 
case of noun phrases). Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.

\begin{table*}
  \centering
  \begin{tabular}{|l|l|l|}
  \hline
  \textbf{Structure} & \textbf{Description} & \textbf{Examples} \\
  \hline
  \texttt{n} & Noun & house \\
  \texttt{adj + n} & Adjective + noun & long road \\
  \texttt{det + n} & Determiner + noun & an apple, the chance \\
  \texttt{num + n} & Numeral + noun & five pieces \\
  \texttt{num + adj + n} & Numeral + adjective + noun & five big pieces \\
  \texttt{det + num + adj + n} & Determiner + numeral + adjective + noun & the five big pieces \\
  \hline
  \end{tabular}
  \caption{Example of the most common structures of noun phrases}
  \label{table:nps}
\end{table*}

\begin{savenotes}
\begin{table*}
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{Phrase type} & \textbf{Description} & \textbf{Examples} & \textbf{Chunk tags} \\
    \hline
    \multirow{5}{*}{\texttt{NP}} & \multirow{5}{*}{Noun phrase}    & noun & number*, person, \\
                                 &                                 & determiner-noun & possessor*, case\\
                                 &                                 & numeral-noun & \\
                                 &                                 & adjective-noun & \\
                                 &                                 & determiner-adjective-noun & \\
    \hline
    \multirow{5}{*}{\texttt{VP}} & \multirow{5}{*}{Verb phrase}    & finite-verb & number, person, \\
                                 &                                 & have-participle & tense/mood, possessive*, \\
                                 &                                 & be-gerund & negation* \\
                                 &                                 & must-infinitive & \\ 
                                 &                                 & want\_to-infinitive & \\ 
    \hline
    \multirow{5}{*}{\texttt{PP}} & \multirow{5}{*}{Postpositional phrase} & preposition-noun & number, person, \\
                                 &                                        & preposition-determiner-noun & possessor*, case \\
                                 &                                        & preposition-numeral-noun & \\
                                 &                                        & preposition-adjective-noun & \\
                                 &                                        & preposition-determiner-adjective-noun & \\
    \hline
    \multirow{2}{*}{\texttt{GenP}} & \multirow{2}{*}{Genitive phrase}  & As with \texttt{PP} except & number, person, \\
                                   &                                   & with the preposition `of'. & possessor*, case \\
    \hline
    \multirow{2}{*}{\texttt{AdjP}} & \multirow{2}{*}{Adjective phrase}  & adjective & n/a \\
                                   &                                    & ``more''--adjective & \\
    \hline
    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor*,\footnote{Treated as a noun phrase with an implied noun (``the largest [book]'')} case \\
                                   &                                     & the--``most''--adjective & \\ 
    \hline
%    \texttt{Q\_mark} & Question mark  & & \\
%    \hline
  \end{tabular}
  \caption{Phrase types in the English-Kazakh chunker}
  \label{table:phrases} 
\end{table*}
\end{savenotes}

As it is shown above, the most probable sequence of world classes are built phrases and 
for English-Kazakh chunker there are used next kind of phrases:

%% EXAMPLES ?

\begin{description}
\item[Noun phrases:] Consider the following example: the chunker identifies the English sequence 
  the large book  (determiner–adjective–noun) as a noun-phrase chunk. It translates into 
  Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), 
  possessor (to be determined, as the noun кітап ('book') could receive a 3rd-person possesive 
  ending (кітабы) later if the context were, for instance, the large book of animals, аңдардың үлкен кітабы), 
  and case (to be determined as it could be, for instance, accusative in I saw the large book, Мен үлкен кітапты көрдім ). Also 
  noun form of the verb, for instance,  gerund form is determined as NP phrase (Her playing is good – Оның ойыны жақсы).
\item[Prepositional phrases:] English prepositional phrases are translated into Kazakh as postpositional phrases, 
  there are three possible outcomes with different cases:
   \begin{itemize}
    \item Genitive -{N}{I}ң,\footnote{Upper case letters in braces (such as {N}) represent hypothetical archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied. For instance, in the genitive ending “-{N}{I}ң”, the archiphoneme {N} may be realised as т, д, or н and the archiphoneme {I} may be realised as  і  or ы depending on the previous phonological context. } which will be marked GenP:  
        [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
    \item Translating the locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
        phonological context.} ablative ``-{D}{A}н'', etc., except the genitive ``-{N}{I}ң'' cases: 
        [PP [P from ] [NP the five cars] ] $\rightarrow$ [PP [NP бес көлік] [P -тен ] ]
    \item With nouns such as аст `under', үст `on', etc. which takes possessive from main noun:  
        [PP [P under ] [NP the garden] ] $\rightarrow$ [PP [NP [GenP [NP бақша ][P -ның]] [NP астын]] [P -да ] ]
    \end{itemize}
\item[Verb phrases:] Translation verb phrases into Kazakh  are not always be straightforward, for instance: 
  present simple and future are rendered using the same tense in Kazakh (I play $\rightarrow$ Мен ойнаймын;  I 
  will play $\rightarrow$ Мен ойнаймын), this ambiguity problem may be solved by lexical selection, 
  which will choose right tense by watching context; tenses expressing continued activity, such as the 
  English present continuous or past continuous (I am playing, I was playing), have to be detected and 
  mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is 
  found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary 
  verb (жатыр, отыр) is used to express number and person agreement,\footnote{Actually, Kazakh language uses four auxiliary verbs: жатыр ('lie', 
      used when the activity takes a long time), отыр ('sit', used when the activity appears to be done in a sitting position), тұр ('stand', 
      when the takes a short time), and жүр (when the activity repeats regularly). Choosing the most adequate auxiliary verb is 
      hard without a semantic analysis, which is not easily available in Apertium. Our current choice (an approximation) 
      is жатыр ('lie') for the present continuous and отыр ('sit') for the past continuous.} 
  modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), 
  and a form of the copula (absent in present tense); the subject receives the genitive or dative case; negative 
  verbs for present continuous and past tense are translated by adding negative word ``емес'' for present 
  perfect continuous (He has not been playing -- Ол ойнап отырған емес) and  ``жоқ'' for past tense and 
  present continuous (see examples above).
\item[Adjectival phrases:] In Kazakh noun phrases, adjectives come before nouns and do not show any agreement 
  with nouns.  Adjectives can also appear in separate adjective phrases. Here are some examples:
\begin{itemize}
\item The adjective alone, marked AdjP: [AdjP big ] $\rightarrow$ [AdjP үлкен] 
\item Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation 
   chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  more  interesting ]  $\rightarrow$ [AdjP қызығырақ]
\item For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation 
   is built using ``ең'' + adjective:  [SupP  the biggest ]  $\rightarrow$ [SupP ең үлкен] 
   [SupP  the most important]   $\rightarrow$ [SupP ең маңызды]
\end{itemize}
Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when 
modified by a genitive phrase: the most beautiful of people $\rightarrow$ адамдардың ең әдемісі); one could 
say that they are treated as NPs with an implied noun.
\end{description}

\begin{table*}
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \multicolumn{4}{|c|}{\textbf{Affirmative forms}} \\
    \hline
    \textbf{English form} & \textbf{Example} & \textbf{Morphemes} & \textbf{Translation} \\
    \hline
    \multirow{2}{*}{Present simple} & \multirow{2}{*}{I play} & \multirow{2}{*}{ойна + <aorist> + <pers.>} & Мен ойнаймын \\
                                    &                         &                                            & I play \\
    \hline
    \multirow{2}{*}{Present continuous} & \multirow{2}{*}{I am playing} & ойна + <perf. part.>  & Мен ойнап жатырмын \\
                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lie \\
    \hline
    \multirow{2}{*}{Past continuous} & \multirow{2}{*}{I was playing} & ойна + <perf. part.>  & Мен ойнап отырдым \\
                                    &                         & отыр + д{I} <past>  + <pers.>     & I playing sat \\
    \hline
    \multirow{2}{*}{Obligation} & \multirow{2}{*}{I must go} & Мен  + <gen.>  & Менің баруым керек \\
                                    &                         & бар + <ger.> +  <px1sg> + керек    & My going necessary [is] \\
    \hline
    \multirow{2}{*}{Should} & \multirow{2}{*}{I should go} & Мен  + <gen.>  & Менің Менің барғаным жөн \\
                                    &                         & бар + <past.ger.> +  <px1sg> + жөн & My was.going proper [is] \\
    \hline
    \multicolumn{4}{|c|}{\textbf{Negative forms}} \\
    \hline
    \multirow{2}{*}{Present cont.} & \multirow{2}{*}{I am not playing} & ойна + <perf. part.>  & Мен ойнап жатқан жоқпын \\
                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lying not \\
    \hline
    \multirow{2}{*}{Ability} & \multirow{2}{*}{I can not play} & ойна + <impf. part.>  & Мен ойнай алмаймын \\
                                    &                         &   ал + <neg.> + <aor.> + <p1>     & I playing can.not \\
    \hline
    \multirow{3}{*}{Present perf. cont. } & \multirow{3}{*}{You have not been playing} & ойна + <perf. part.>  & Сіз ойнап отырған емессі \\
                                    &                         & +  отыр + <past>  & You playing sat not  \\
                                    &                         & + eмес <cop.neg> + <p2> & \\
    \hline

  \end{tabular}
  \caption{Examples of Verb phrase translations}
  \label{table:vps}
\end{table*}

\subsubsection{Interchunk processing}

The second round of structural transfer is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
\begin{itemize}
\item Inter-chunk agreement between subject noun phrase and verb phrase in their person and number.
\item Assigning case to noun phrases (which are generated without case by the chunker): for instance, 
  accusative case for objects (I bought the table $\rightarrow$ Мен үстелді  сатып алдым), genitive case for 
  obligatory constructs (I must see $\rightarrow$ менің көруім керек),  dative case for the verb to 
  need (I need a doctor $\rightarrow$Маған дәрігер керек), locative case for 
  possession (I have a book $\rightarrow$ Менде кітап бар), etc.
\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] $\rightarrow$ Мен[1] үстелді[3] сатып алдым[2]), 
  placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
\item Adding question word ма/ме/ба/бе in the end of question, if question mark chunk ``?'' is detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).
\end{itemize}

The set of rules has to be extended, as many combinations of the above phenomena are still not covered.

\subsubsection{Postchunk}

Additional cleaning stage was created for removing extra tags which did not determine through the chunker and 
interchunk steps, for instance, noun was not acting as object and did not get accusative case, so not determined 
tag was \texttt{<CD>} after ``cleaning'' it will be set on nominative case; at this step, also carried out 
agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singular>), vowel and consonants 
agreement between last word of question and question particle (Сіз келдіңіз + ме? - Сіз келдіңіз + бе?).

\section{Evaluation and results}

Current system can translate simple sentences and questions. The English-Kazakh bilingual dictionary 
contains 13,135 entries, ``chunker'' rules are 168 and interchunk rules are 148. 
Evaluating was tested against  revision 58998 in the Apertium SVN1. Lexical coverage is calculated 
over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014} 
Table~\ref{table:coverage} presents the size of the corpora and the vocabulary coverage of the system for the given corpus.

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|}
    \hline
    \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) \\
    \hline
    SETimes & 5.1 mil. & 97.90 \\
    NewsCommentary & 6.5 mil. & 96.27 \\
    EuroParl & 54.5 mil. & 97.95 \\
    Wikipedia & 1.8 bil. & 84.67 \\
    \hline
  \end{tabular}
  \caption{Vocabulary coverage of the English--Kazakh system over four available corpora.}
  \label{table:coverage}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|}
    \hline
    \textbf{System} & \textbf{BLEU} & \textbf{WER} \\
    \hline
    \texttt{apertium-eng-kaz} & 44.23 & 42.88 \\
    Google & 62.00 & 25.92 \\
    Sanasoft & 21.00 & 74.52 \\
    \hline
  \end{tabular}
  \caption{Metric results for the three systems compared.}
  \label{table:metrics}
\end{table}

Evaluating the output of the system was measured with BLEU \citep{papineni02} and Word Error Rate 
metric \citep{levenshtein/1966}. We chose a short text in English, and by postediting output 
of English-Kazakh system, was build parallel text. The output of each of the machine
translation systems was posteditted independently to avoid biasing one particular system.

As can be seen,  coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
available RBMT system, but fall short of the state-of-the-art performance. However, unlike the state of 
the art, which uses unavailable corpora, our system is free and open and can be improved.

MT systems have some mistakes in translations, which was shown in <...> brackets. Google MT system 
has a higher BLEU score, but in translation separate phrases, it has common error: it does not assign 
the right possessive and case. Sanasoft system has more errors with choosing not right translations 
and out of vocabulary words.

\begin{table*}
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline 
    \textbf{Structure} & \textbf{English} & \textbf{System} & \textbf{Translation} \\
    \hline 
    \multirow{6}{*}{Noun phrases} & \multirow{3}{*}{My difficult exercises} & Apertium & Менің қиын жаттығуларым \\
                                  &                                         & Sanasoft & Менің қиын жаттықтырып *жатыр \\ % <not right POS>
                                  &                                         & Google   & Менің қиын *жаттығулар \\ % <no 1 person possessive>
                                  & \multirow{3}{*}{Conan Doyle}            & Apertium & *Дойлдың Конан \\
                                  &                                         & Sanasoft & *Conan *Doyle \\ % <not right POS>
                                  &                                         & Google   & Конан Дойл \\ % <no 1 person possessive>
    \hline 
    \multirow{3}{*}{Adpositional phrases}  & \multirow{3}{*}{Under three big trees}  & Apertium & үш үлкен ағаштың астында \\
                                           &                                         & Sanasoft & *Three үлкен *ағаштар астында \\ % <unk> <no genitive>
                                           &                                         & Google   & үш үлкен *ағаштарды астында \\ % <no genitive>
    \hline 
    \multirow{3}{*}{Adjective phrases}    & \multirow{3}{*}{The most beautiful}      & Apertium & ең әдемі \\ 
                                          &                                          & Sanasoft & *Көпшілік әдемі \\ % <not right translation> 
                                          &                                          & Google   & ең әдемі \\
    \hline
    \multirow{6}{*}{Modal verbs}  & \multirow{3}{*}{I must pay}             & Apertium & Мен жүргізе аламын \\
                                  &                                         & Sanasoft & Мен must pay \\ % <no translation>
                                  &                                         & Google   & *Мен *төлеуі тиіс \\ % <no genitive> <no possessive>
                                  & \multirow{3}{*}{It must be James}       & Apertium & Бұл Джеймс *болады \\ % <should be "болуы керек"> 
                                  &                                         & Sanasoft & *Оған Джеймс *must *болып *жатыр  \\ % <no translation>
                                  &                                         & Google   & Джеймс болуы тиіс \\ % <no genitive> <no possessive>
    \hline 
    \multirow{3}{*}{Questions}    & \multirow{3}{*}{Is it right?}           & Apertium & *Екен *дұрыстың ол? \\ % extra word + extra case 
                                  &                                         & Sanasoft & Бұл *түзуi? \\ % <not right translation>
                                  &                                         & Google   & Бұл дұрыс па? \\
    \hline 
  \end{tabular}

  \caption{Qualitative evaluation}
  \label{table:qualeval}
\end{table*}

\section{Conclusions}

We have presented the rule-based MT system from English to Kazakh. The current English-Kazakh machine translation 
system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and 
contains a good vocabulary for testing purposes. 
We plan to continue developing English-Kazakh pair, going to extend the coverage to 98\% on the corpora, 
which we used before. Additionally, we will improve the quality of translation by adding more rules, such like 
constraint grammar rules, structural transfer and lexical rules. The future plan is to use the created data with other 
open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh. 
Our system is available as free/open-source software and the whole system may be downloaded 
from SourceForge.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/incubator/apertium-eng-kaz}}

\section*{Acknowledgements}

MLF thanks the Kazakh state program for the attraction of foreign scholars and Prof. Ualsher Tukeyev for supporting his visit to the Kazakh National University, where part of this work was carried out. We also thank Prof. Tukeyev for his valuable input.

\bibliographystyle{apalike}

\bibliography{2015-eamt-engkaz}

\end{document}