### 2015-03-24 2015-nodalida-ruscg/ruscg.tex
added ambiguity intro
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59500 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File nodalida2015.tex
%
% Contact beata.megyesi@lingfil.uu.se
%
% Based on the instruction file for EACL 2014
% which in turn was based on the instruction files for previous 
% ACL and EACL conferences.

\documentclass[11pt]{article}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian,english]{babel}

\usepackage{times}
\usepackage{latexsym}
\usepackage{fixltx2e} %allows subscripts
%\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage[small,bf]{caption}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

\usepackage{nodalida2015}

\usepackage{linguex}
\usepackage{needspace}

\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}

\newcommand{\ft}[1]{\marginpar{\scriptsize F: #1}} % Fran's comments
\newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments

\title{A preliminary constraint grammar for Russian}

\author{Francis M. Tyers \\
  HSL-fakultehta, \\
  UiT Norgga árktalaš universitehta, \\
  N-9018 Romsa \\
  {\tt francis.tyers@uit.no} \\\And
  Robert Reynolds \\
  HSL-fakultehta, \\
  UiT Norgga árktalaš universitehta, \\
  N-9018 Romsa \\
  {\tt robert.reynolds@uit.no} \\}

% \author{Author1 \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   {\tt email@domain.com} \\\And
%   Author2 \\
%   Affiliation / Address line 1 \\
%   Affiliation / Address line 2 \\
%   Affiliation / Address line 3 \\
%   {\tt email@domain.com} \\}
% 
\date{2015}

\begin{document}
\maketitle
\begin{abstract}
 This paper presents preliminary work on a constraint
 grammar based disambiguator for Russian. Russian is
 a Slavic language with a high degree of both in-category
 and out-category homonymy in the inflectional system.
 The pipeline consists of a finite-state morphological
 analyser and constraint grammar. The constraint 
 grammar is tuned to be high recall at the expense of 
 low precision.
\end{abstract}

\section{Introduction}

Our purposes: Processing stressed wordforms and open-source machine translation

\cite{Karlsson-90}

% objectives:
%% make a free/open-source morphology for russian based on Z's dictionary that includes stress

\section{Review of literature}

State-of-the-art morphological analysis in Russian is primarily based on
finite-state technology
\cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by 
\newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale 
morphological transducers of Russian are based on the
forward-looking \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
This dictionary gives fine-grained morphological specifications for more than
100~000 words, including inflectional endings, morphophonemic alternations, 
stress patterns, exceptions, and idiosyncratic collocations.
We base our work on the recently developed [anonymous] morphological transducer, which is also based on Zaliznjak's dictionary (Author, 2015).\footnote{The [anonymous] transducer is implemented using a two-level 
morphology \cite{Koskenniemi-84}, and can be compiled using either {\tt xfst}
\cite{Beesley.Karttunen-03} or {\tt hfst} \cite{hfst-11}} This finite-state transducer (FST) generates all possible 
morphosyntactic readings of each wordform, regardless of its frequency or probability. Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below. Because of this, one of the foundational steps in Russian natural language processing is morphosyntactic disambiguation, or \emph{snjatie omonimii}.

% FRAN
% different CGs 
% faroese
% estonian

 -- and our Constraint 
Grammar\footnote{Implemented using vislcg3 constraint grammar parser
(http://beta.visl.sdu.dk/cg3.html).}
\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
some readings based on syntactic context.

\cite{trosterud2009}

\section{Ambiguity in Russian} \label{ambiguity}

We identify three different types of morphosyntactic ambiguity: 
intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
congruent. A recent study by Author (2015) found that about 7.5\% of 
morphosyntactic ambiguity in a corpus of Russian resulted in stress placement 
ambiguity. The following examples make use of stress ambiguity to illustrate 
each kind of ambiguity. \emph{Intraparadigmatic} ambiguity refers to homographic 
wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.

\needspace{4\baselineskip} % keeps example all together (no orphan line) 
\ex. Intraparadigmatic homographs \label{ex:intrahom}
\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
    \label{ex:bodySGGEN}
\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
    \label{ex:bodyPLNOM}

The remaining two types of ambiguity occur between lexemes. 
\emph{Morphosyntactically incongruent} ambiguity occurs between homographs that 
belong to separate lexemes, and whose morphosyntactic values are different, as 
shown in \ref{ex:MSincongruent}.

\needspace{4\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
\a. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
\b. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'

\emph{Morphosyntactically congruent} ambiguity occurs between homographs 
that belong to separate lexemes, and whose morphosyntactic values are identical, 
as shown in \ref{ex:MScongruent}. As indicated in Table~\ref{table:ambiguity}, 
this kind of ambiguity is relatively rare, and resolving this ambiguity requires 
the use of technologies such as Word Sense Disambiguation.

\needspace{6\baselineskip} % keeps example all together (no orphan line) 
\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
\c. ...\\
	...

Table \ref{table:ambiguity} shows the proportional prevalence of each kind of 
ambiguity. Note that these proportions do not sum to 1, since a given token may 
exhibit more than one kind of ambiguity. The first column shows the proportion 
of all tokens in a corpus that have each kind of ambiguity. The second column 
shows what proportion of ambiguous tokens exhibit each kind of ambiguity.

Table~\ref{table:ambiguity} shows that most morphosyntactic ambiguity in 
unrestricted Russian text is rooted in intraparadigmatic and morphosyntactically 
incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
analysis can help disambiguate these forms. Morphosyntactically congruent stress 
ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
and since congruent stress ambiguity represents only a small percentage of 
ambiguous wordforms, we leave it to future work.

\begin{table}
  \centering
  \begin{tabular}{l|rr}
    \hline
    \textbf{Type}  & \textbf{all tokens} & \textbf{ambig. tokens} \\
    \hline
    Intraparadigm. & 0.590                   & 0.909   \\
    Incongruent    & 0.277                   & 0.427   \\ 
    Congruent      & 0.120                   & 0.180    \\ 
    \hline
  \end{tabular}
  \caption{Proportional frequency of different types of morphosyntactic ambiguity in unrestricted text}
  \label{table:ambiguity}
\end{table}

% numbers not perfect, e.g. лицо should be 'cong.' but will turn up as 'intra.'

\section{Pipeline}

% FRAN

\subsection{Morphological analyser}
% ROB

\cite{yablonsky1999russian}

% det/prn (e.g. pos... их, я, мой)
% 'over'generation (e.g. all dets get all gen/nbr/case combinations)

\subsection{Disambiguation rules}
%FRAN

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \hline
         & {\sc select} & {\sc remove} & {\sc map} \\
    Safe & & & \\
    Safe heuristic  & & & \\
    Heuristic &  & & \\
    Syntax labelling &  &   & \\ 
    \hline
  \end{tabular}
  \caption{Rules in the grammar are separated into four sections. }
\end{table}

\section{Development process}
%FRAN

\subsection{Combining with a statistical tagger}

\section{Evaluation}

\subsection{Intrinsic}
%FRAN

\begin{table*}
  \centering
  \begin{tabular}{|l|r|r|r|r|r|}
    \hline
    \textbf{Domain} & \textbf{Tokens} & \textbf{Precision} & \textbf{Recall} & \textbf{F-score} & \textbf{Ambig. solved} \\
    \hline
    Wikipedia       & 7,857      & 0.506        & 0.996    & 0.671 & 44.92\%  \\
    Literature      & 1,652      & 0.473        & 0.984    & 0.638 & 42.95\%  \\
    News            & 642        & 0.471        & 0.990    & 0.638 & 41.60\%  \\
    \hline
    \textbf{Average}& 10,150     & 0.498        &  0.994   & 0.663 & 44.39\% \\
    \hline
  \end{tabular}
  \caption{Results for the test corpora.}
\end{table*}

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|}

    \hline
    Error in original        &   & 2 \\
    Missing lexeme           &   & 8  \\          % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
    Partial lexeme           &   & 2  \\          % франко-норманнское, рельс
    Missing analysis         &   &    \\        % 
    Equally valid analysis   &   &    \\   % большой<adj><comp><pred> ['больше<adv>'] (?)
    Correct reading removed  &   & 34   \\ % 
    \hline
                             &   & 46 \\
    \hline
  \end{tabular}
  \caption{Error analysis of false negatives}
\end{table}

% unknown words from test corpus, categorisation

% rule error analysis

\subsection{Extrinsic}
% ROB accent placement task

\section{Future work}

% ROB
% FRAN
% func analysis
% dep analysis

\section{Conclusions}

%\section*{Acknowledgments}
%
%Do not number the acknowledgment section. Do not include this section
%when submitting your paper for review.

\bibliographystyle{acl}
\bibliography{ruscg}

\end{document}