### 2015-05-06 2015-nodalida-ruscg/presentation_ruscg.tex
adding presentation
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60274 72bbbca6-d526-0410-a7d9-f06f51895060

%% Hey, emacs(1), this is -*- mode: latex; coding: utf-8; fill-column: 79 -*-

\documentclass{beamer}
\usepackage[T2A,T1]{fontenc}
\usepackage[russian,english]{babel}
\usepackage{relsize}
\usepackage{linguex}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\usetheme{Montpellier}
\usecolortheme{beaver}

\title[A preliminary constraint grammar for Russian] % (optional, only for long titles)
{A preliminary constraint grammar for Russian}
%\subtitle{}
\author[Tyers, Reynolds] % (optional, for multiple authors)
{F.~Tyers \and R.~Reynolds}
\institute[University of Troms\o] % (optional)
{
  Department of Languages and Linguistics\\
  University of Troms\o
}
\date[NODALIDA-CG 2015] % (optional)
{Constraint Grammar Workshop - Methods, Tools and Applications (NODALIDA 2015)}
\subject{Constraint Grammar, Russian, Morphosyntactic disambiguation}

\begin{document}
\frame{\titlepage}

\section{Outline}

\begin{frame}
\frametitle{Presentation outline}
%\framesubtitle{}
\begin{itemize}
	\item Ambiguity in Russian
	\pause
	\item Finite-state transducer for Russian part-of-speech tagging
	\pause
	\item Constraint Grammar disambiguation rules
	\pause
	\item Development process
	\pause
	\item Evaluation
	\pause
	\begin{itemize}
		\item Qualitative
		\pause
		\item Task-based (word stress placement)
		\pause
		\item Combining with HMM tagger
	\end{itemize}
	\pause
	\item Future work
\end{itemize}
\end{frame}

\section{Ambiguity in Russian} %ROB
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Finite-state transducer for Russian part-of-speech tagging} %ROB
\begin{frame}
\frametitle{New finite-state tools for Russian morphological analysis}
\framesubtitle{}
\begin{itemize}
	\item We developed a finite-state transducer using the two-level formalism \cite{Koskenniemi-83}
	\pause
	\begin{itemize}
		\item can be used to analyze and generate Russian wordforms, with or without stress marks
		\pause
	\end{itemize}
	\item Since Russian has systematic syncretism and widespread homonymy, we are developing a constraint grammar \cite{Karlsson.Voutilainen.ea-95} to disambiguate such cases
	\pause
	\item implemented in the freely available CG3 system (http://beta.visl.sdu.dk/cg3.html)
\end{itemize}
\end{frame}

\section{Constraint Grammar disambiguation rules} %FRAN

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Development process} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Evaluation}
\subsection{Qualitative evaluation} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\subsection{Task-based evaluation} %ROB

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\subsection{Combining with HMM tagger} %FRAN

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Conclusions and future work} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\bibliographystyle{apalike}
\begin{tiny}
\bibliography{ruscg}
\end{tiny}
\end{document}