### 2015-06-26 2015-compmorphon-morphver/poster.tex
draft poster, including Makefile and logo graphics
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61089 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[fontscale=0.3,landscape,paperwidth=48in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm

\usepackage{setspace}
\usepackage{multicol}
\usepackage{multirow}

\usepackage{tikz}
%\usepackage{pgfbaselayers}
%\pgfdeclarelayer{background}
%\pgfdeclarelayer{foreground}
%\pgfsetlayers{background,main,foreground}

%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\selectcolormodel{cmyk}

\definecolor{bordercol}{RGB}{0,0,0}
%\definecolor{headercolone}{RGB}{6,130,22}
\definecolor{headercol}{cmyk}{1,0.64,0,0.6}
\definecolor{headercolone}{cmyk}{1,0.64,0,0.6}
\definecolor{headercoltwo}{cmyk}{0.44,0,0.15,0.17}
%\definecolor{boxcolor}{RGB}{176,208,223}
\definecolor{boxcolor}{cmyk}{0.3,0.08,0.08,0.1}
\definecolor{headerfontcol}{RGB}{255,255,255}
\newcommand{\hilitetwo}[1]{{\addfontfeature{Color=99333399}#1}}
\newcommand{\hiliteone}[1]{{\addfontfeature{Color=06821699}#1}}
\newcommand{\hilitegrey}[1]{{\addfontfeature{Color=77777799}#1}}

%%% Utility functions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Save space in lists. Use this after the opening of the list %%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
        \setlength{\itemsep}{1pt}
        \setlength{\parskip}{0pt}
        \setlength{\parsep}{0pt}
}

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}

\usepackage{expex}

\usepackage{fontspec}
\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}
\newfontfeature{IPA}{+mgrk}
%\setromanfont[IPA]{FreeSerif}
%\setromanfont[IPA,Scale=0.8]{CMU Serif}
\setromanfont[IPA]{Liberation Serif}
\setmonofont[IPA]{DejaVu Sans Mono}
\usepackage[small,bf]{caption}
\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
\newfontfamily\qgmk[IPA,Scale=0.8]{CMU Serif}
\newfontfamily\litamono[IPA,Scale=0.5]{DejaVu Sans Mono}

\newcommand{\tilda}{{\qipa ∼}}

\newcommand{\tags}[1]{\hiliteone{#1}}
\newcommand{\tag}[1]{\hilitegrey{<}\hiliteone{#1}\hilitegrey{>}}
\newcommand{\archiphon}[1]{\hilitegrey{\{}\hiliteone{#1}\hilitegrey{\}}}

%\newfontfamily\htwo[IPA,Scale=1.2]{FreeSans}}
%\newfontfamily\htwofont[IPA,Scale=1]{CMU Sans Serif}
%\newfontfamily\htwofont[IPA,Scale=1,Color=333344FF]{CMU Sans Serif}
\newfontfamily\htwofont[IPA,Scale=1,Color=111111FF]{CMU Sans Serif}
%\newfontfamily\titlefont[IPA,Scale=0.52]{CMU Serif}
\newfontfamily\titlefont[IPA,Scale=0.7]{CMU Serif}
%\newfontfamily\titlefont[IPA,Scale=0.55]{DejaVu Serif}

%\newcommand{\htwo}[1]{::: {\htwofont #1} :::}%\hrule} %\hline\\}
%\newcommand{\htwo}[1]{{\htwofont \textbf{:::#1:::}}}%\hrule} %\hline\\}
\newcommand{\htwo}[1]{{\htwofont \textbf{\dotfill{}#1\dotfill{}}}}

\usepackage{graphicx}  % [dvipdfm]

%\definecolor{MyGray}{rgb}{0.96,0.97,0.98}
%\newenvironment{codebox}{%
%   \begin{lrbox}{\@tempboxa}\begin{minipage}{\columnwidth}}{\end{minipage}\end{lrbox}%
%   \colorbox{MyGray}{\usebox{\@tempboxa}}
%}
%\newcommand{\codeex}[1]{\begin{codebox}#1\end{codebox}}

%\definecolor{grey}{rgb}{0.96,0.97,0.98}
\definecolor{grey}{rgb}{0.91,0.91,0.91}
\newcommand{\codeex}[1]{
   \fbox{\colorbox{grey}{
         \begin{minipage}[t]{0.91\textwidth}{\litamono
            #1
         }\end{minipage}
      }
   }
}

\newcommand{\blank}[1]{\underline{\hspace{#1}}}

\usepackage{natbib}

\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}

\usepackage{subfigure}
\usepackage{booktabs}

%%\bibpunct{(}{)}{;}{A}{,}{,}
%\bibdata{paper}

\newcommand{\citemultileft}[1]{(\citeauthor{#1}, \citeyear{#1}}
\newcommand{\citemultimid}[1]{\citeauthor{#1}, \citeyear{#1}}
\newcommand{\citemultiright}[1]{\citeauthor{#1}, \citeyear{#1})}
\newcommand{\citetwoyears}[2]{\citeauthor{#1} (\citeyear{#1} and \citeyear{#2})}

% for glosses
\newcommand{\eng}[1]{`{\em #1}'}
%dammit, sc doesn't seem to be working
\newcommand{\gmk}[1]{{\qgmk \textsc{#1}}}

\usepackage{enumitem}
\setlist{nolistsep,leftmargin=*}
\newenvironment{itemise}[1]{
        \begin{itemize}\setlength{\leftmargin}{-4em}\setlength{\itemsep}{-0.2em}
        \vspace{-0.5em}
        #1
}{
        \end{itemize}
        \vspace{-2pt}
}

%\newcommand{\h2}[1]{{\big

\begin{document}
	% To get it to be the same size consistently on all machines..
	\special{papersize=48in,36in}
	\setlength{\pdfpageheight}{\paperheight}
	\setlength{\pdfpagewidth}{\paperwidth}

	%%% Setting Background Image %%%
	%\background{}

	\begin{poster}{
		grid=false,
		%eyecatcher=false,
		borderColor=bordercol,
		headerColorOne=headercolone,
		headerColorTwo=headercoltwo,
		%headerColorOne=blue,
		%headerFontColor=white,
		% Only simple background color used, no shading, so boxColorTwo isn't necessary
		boxColorOne=boxcolor,
		headershape=roundedright,
		headerborder=open,
		headerheight=0.1\textheight,
		%headershape=roundedright,
		%headershade=plain,
		%headerfont=\Large\textsf, %Sans Serif
		%headerfont=\Large\sf\bf,
		textborder=rectangle,
		%background=plain,
		%background=user,
		background=none,
		headerborder=open,
		boxshade=plain,
		textborder=roundedleft,
		columns=4,
	}{ %Eye Catcher, empty if option eyecatcher=false - unused
		\includegraphics[height=6.5em]{uitlogo}
		\hspace{2cm}\begin{minipage}[t]{9em}
			\vspace{-2cm}
			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
			\end{minipage}
			\hspace{2cm}\includegraphics[height=6.5em]{iu_tab}
	}{ %Title (centered top)
		{\vspace{0pt}\hspace{1cm}
			{\titlefont \textsc{Finite-state morphologies and text corpora\\as resources for improving morphological descriptions}}}
	}{ %Authors (centered top, below title)
		\vspace{-0.5em}
		\begin{center}
			\begin{minipage}[t]{9em}
				\begin{spacing}{0.4}
				{Francis M.\ Tyers}\\
				{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
				\end{spacing}
			\end{minipage}
			\begin{minipage}[t]{9.5em}
				\begin{spacing}{0.4}
					{Tommi Pirinen\vphantom{y}}\\
					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\\\texttt{FIXME Flammie}}
				\end{spacing}
			\end{minipage}
			\begin{minipage}[t]{12em}
				\begin{spacing}{0.4}
					{Jonathan North Washington}\\
					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}
				\end{spacing}
			\end{minipage}
		\end{center}
	}{ %More eye catchers, on the right side
		\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
	}

	\headerbox{Random box 1}{name=boxone,column=0,row=0}{
		\scalebox{0.78}{
			hargle
		}
	}

	\headerbox{Morphological transducers}{name=morphtrans,below=boxone}{
		\htwo{Morphological transducers}
		\begin{itemize}
			\item Efficient (in speed \& size) models of a language's morphology
			\item Take a surface form, and produce valid lexical form(s)
			\item Take a lexical form, and produce valid surface form(s)\vspace{-0.1ex}\\
		%\end{itemize}
		\scalebox{0.98}{
			алдым \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
		}
		\end{itemize}

		\htwo{Transducers for Turkic languages}
		\begin{itemize}
			\item Turkish (\hiliteone{Çöltekin, 2010 \& 2014}; \hilitetwo{Oflazer, 1994})
			\item Crimean Tatar (\hilitetwo{Altıntaş, 2001})
			\item Turkmen (\hilitetwo{Tantuğ et al., 2006})
			\item Kazakh (\hilitetwo{Бекманова \& Махимов, 2013})
			%\vspace{1pt}\hfill{} \hiliteone{GPL (=free and open)}!
			\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
		\end{itemize}

		\htwo{Framework: HFST}
		\begin{itemize}
			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
		\end{itemize}
		\htwo{Approach}
		\begin{itemize}
			\item morphotactics implemented in \texttt{lexc}
			\item morphophonology implemented in \texttt{twol}
			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}\vspace{-0.1ex}\\
			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
		\end{itemize}

	}

	\headerbox{Categorisation}{name=adjectives,column=2}{
		\htwo{Adjectives}
			\begin{itemize}
				\item morphologically distinct adjective classes
				\item most sources claim:\ adjectives can be used substantively and adverbially
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item but not all adjectives have all of the following:\\
				comparative forms, substantive readings, adverbial readings
				\item Our approach:\ categorisation 
				\item if properly categorised, only correct forms are analysed and generated
			\end{itemize}

			\scalebox{0.7}{
				\begin{tabular}{lllll}
					\toprule
					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
					\midrule
					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
					\bottomrule
				\end{tabular}
			}

		\htwo{Adverbs}
			\begin{itemize}
				\item Certain adverbs have special attributive and ablative forms
				\item Mostly time adverbs
				%\item Mostly time adverbs, e.g.\\
				%{\qipa бүгүн} \eng{today}, {\qipa быйыл} \eng{this year}, {\qipa кечээ} \eng{yesterday}, {\qipa жана} \eng{just now}
				%\item Attributive forms:\\
				%{\qipa бүгүнкү} \eng{today's}, {\qipa быйылкы} \eng{this year's}, {\qipa кечээги} \eng{yesterday's}, {\qipa жанагы} \eng{just now}
				%\item Ablative forms:\\
				%{\qipa бүгүнтөн} \eng{from today}, {\qipa быйылтан} \eng{from this year}, {\qipa кечээтен} \eng{from yesterday}, {\qipa жанатан} \eng{from just now}
				\item Some also have noun readings:\ regular ablative, other cases:
				%{\qipa бүгүндөн} \eng{from today}, {\qipa быйылдан} \eng{from this year}
			\end{itemize}
			\scalebox{0.83}{
			\begin{tabular}{lllll}
				\toprule
				word & {\qipa бүгүн} & {\qipa быйыл} & {\qipa кечээ} &  {\qipa жана}\\
				gloss & \eng{today} & \eng{this year} & \eng{yesterday} & \eng{just now} \\
				\midrule
				\tag{attr} form & {\qipa бүгүнкү} & {\qipa быйылкы} & {\qipa кечээги} &  {\qipa жанагы}\\
				\tag{adv}\tag{abl} form & {\qipa бүгүнтөн} & {\qipa быйылтан} & {\qipa кечээтен} &  {\qipa жанатан}\\
				\tag{n}\tag{abl} form & {\qipa бүгүндөн} & {\qipa быйылдан} & — & — \\
				\bottomrule
			\end{tabular}
			}\vspace{0.2em}
				\codeex{
					\begin{spacing}{0.8}
					\texttt{LEXICON ADV-WITH-KI-ABL\\
\\
ADV ;\\
ADV-KI ;\\
ADV-ABL ;
					}
					\end{spacing}
					\vspace{0em}
				}
	}

	\headerbox{Ongoing and future work}{name=nedostatki,column=3}{
		\begin{itemize}
			\item Disambiguation, more stems, clean up transducers
			\item Machine translation between these languages
			\item Bring other Kypchak transducers to comparable performance:\\
			Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
			\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
		\end{itemize}
	}

	\headerbox{Further information}{name=getting,column=3,below=nedostatki}{
		\begin{itemize}
			\item Part of \textbf{Apertium Turkic} project:\\
			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
			\item \textbf{Source code} available from Apertium's svn repo
			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
			\texttt{apertium-turkic@lists.sourceforge.net}\\
			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
			\item See our papers in LREC proceedings\\
			(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
			\item And feel free to contact the authors any time!
		\end{itemize}
	}

%	\headerbox{References}{name=references,column=2,below=nedostatki}{
%		%\begin{thebibliography}{1}\itemsep=-0.01em
%		\setlength{\baselineskip}{0.4em}
%	}

		\headerbox{Morphophonology}{name=morphophon,column=1}{

			\htwo{Desonorisation (kaz \& kir)}
				\begin{itemize}
					\item {\texttt \{N\}} desonorises to д after a consonant\\
						алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
						сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
					\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
						сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
						кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
				\end{itemize}

				\vspace{-0.5ex}
				\codeex{
					\texttt{"L Desonorisation"\\
						\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\

					\texttt{"N Desonorisation"\\
						\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
				}
			\vspace{0.1em}

			\htwo{Epenthesis}
				\begin{itemize}
					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
					мур\archiphon{у}н → мурун \eng{nose}\\
					мур\archiphon{у}н+\archiphon{I}м → мурдум \eng{my nose}\\
				\end{itemize}
					\vspace{-0.5ex}
\codeex{
			\begin{spacing}{0.9}
{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
  \vspace{-0.5ex}\hspace{12ex}         matched ;}}\vspace{-1.5em}
  			\end{spacing}
}
\vspace{0.1em}
		}

	\end{poster}
\end{document}