### 2015-09-17 2015-turklang-kazdep/slides/trees/minimal.tex
today's work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61928 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[landscape,a5paper,12pt]{minimal}
\usepackage{tikz-dependency}

\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}
\usepackage[margin=0.1in,landscape]{geometry}
\geometry{bmargin=0in}

\setromanfont{Times New Roman}

\newfontfamily\ll[]{Linux Libertine O}
%\newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
\newfontfamily\udfont[Scale=1.3,Color=FFFFFF99]{Alegreya Sans SC}
\newfontfamily\gmkfont[Scale=1,Color=FFFFFF99]{Alegreya Sans SC}
\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps,Color=FFFFFF99]{Linux Libertine O}

\newcommand{\white}[1]{{\addfontfeature{Color=FFFFFF99}#1}}

\newcommand{\gmk}[1]{{\gmkfont #1}}
\newcommand{\tgloss}[1]{{\em #1}}
\newcommand{\tag}[1]{{\ll \textsc{‹#1›}}}
\newcommand{\udtag}[1]{{\ll \textsc{#1}}}
\newcommand{\udlabel}[1]{{\udfont #1}}

\newcommand{\scriptsize}[1]{#1}

\begin{document}

\resizebox{\textwidth}{!}{%
	\centering
		\white{
		\begin{dependency}[theme = simple, edge style={white!60!white,very thick}] %, font = \small]
			\begin{deptext}[column sep=0.2cm]
				¶
		\end{dependency}}
}
\end{document}